---
title: "Canadian Goose Spatial Location"
author: "Karthiga"
date: "2024-04-04"
output: html_document
---

#### **Index**

#### ** Relationship with Elevation**

#### **1. Loading Data**
#### **2. Removing NA**
#### **3. Method to get BC Albers**
#### **4. Adding x and y to the datasets from BC Albers**
#### **5. Testing if the thinned data is good representation of original data**
#### **6. Conversion to ppp**
#### **7. visualizing Canadian Goose Locations along with elevation data in 2D and 3D**
#### **8. Calculating and Visulaizing number of goose at each level of elevation**
#### **9. Testing homogeneity**
#### **10. calculating the intensity of Canadian goose locations with respect to elevation**
#### **11. calculating the correlation of Canadian goose locations with respect to elevation**
#### **12. calculating presence of clustering**

#### ** Model Fitting**

#### **13. Elevation only Linear Model**
#### **14. Elevation only non-linear model**
#### **15. Elevation only Linear Model vs Elevation only non-linear model**
#### **16. Checking if data deviate significantly from the Elevation only non-linear model prediction**
#### **17. Visualising Residual Elevation only non-linear model prediction**
#### **18. Correlation between co-variates**
#### **19. Checking all rho values of co-variates**
#### **20. Modelling with all co-variates**
#### **21. Comparing All Covariate model with null model and Elevation only Model**
#### **22. Partial Residual of all co-variate model**
#### **23. Fitting Spline Model**
#### **24. Comparing smooth vs all- covariate model**
#### **25. Calculating partial residual of spline model**



```{r}
library(spatstat)
library(maptools)
library(sp)
library(sf)
```
### 1. Loading Data
```{r}
#EXP
#colSums(is.na(c_g_data_5k))

#Loading Data
c_g_data_full = read.csv("C:/Users/karth/Documents/Assignments/block-6/589/project/Samples_CG/canadian_goose.csv",sep = ",")
c_g_data_5k = read.csv("C:/Users/karth/Documents/Assignments/block-6/589/project/Samples_CG/cg_5k_samples.csv",sep = ",")
c_g_data_8K = read.csv("C:/Users/karth/Documents/Assignments/block-6/589/project/Samples_CG/cg_8k_samples.csv",sep = ",")
c_g_data_10k = read.csv("C:/Users/karth/Documents/Assignments/block-6/589/project/Samples_CG/cg_10k_samples.csv",sep = ",")

load("BC_Covariates.Rda")
```

### 2. Removing NA
```{r}
c_g_data_5k<-na.omit(c_g_data_5k)
c_g_data_8K<-na.omit(c_g_data_8K)
c_g_data_10k<-na.omit(c_g_data_10k)
```

### 3. Method to get BC Albers
```{r}
#' Helps to get BC Albers projection
#'
#' @param df 
#'
#' @return BC Albers projection
#'
#' @examples
get_bc_albers <- function(df)
{
 sf_points <- st_as_sf(df, coords = c("decimalLongitude","decimalLatitude"), crs = 4326)
 bc_albers_proj <- "+proj=aea +lat_0=45 +lon_0=-126 +lat_1=50 +lat_2=58.5 +x_0=1000000 +y_0=0 +datum=NAD83 +units=m +no_defs"
 sf_points_bc_albers <- st_transform(sf_points, crs = bc_albers_proj)
 return(sf_points_bc_albers)
}
```

### 4. adding x and y to the datasets
```{r}
sf_points_bc_albers <- get_bc_albers(c_g_data_5k)
c_g_data_5k$x <- st_coordinates(sf_points_bc_albers)[, 1]
c_g_data_5k$y <- st_coordinates(sf_points_bc_albers)[, 2]

sf_points_bc_albers <- get_bc_albers(c_g_data_8K)
c_g_data_8K$x <- st_coordinates(sf_points_bc_albers)[, 1]
c_g_data_8K$y <- st_coordinates(sf_points_bc_albers)[, 2]

sf_points_bc_albers <- get_bc_albers(c_g_data_10k)
c_g_data_10k$x <- st_coordinates(sf_points_bc_albers)[, 1]
c_g_data_10k$y <- st_coordinates(sf_points_bc_albers)[, 2]

print(head(c_g_data_5k))
print(head(c_g_data_8K))
print(head(c_g_data_10k))
```
### 5. Testing if the thinned data is good representation of original data
#### 5.1. Getting rho value of datasets with different numbber of rows
```{r}
library(maptools)
cg_ppp_5 <-ppp(x = c_g_data_5k$x, y = c_g_data_5k$y, window=as.owin(DATA$Window))
cg_ppp_8 <-ppp(x = c_g_data_8K$x, y = c_g_data_8K$y, window=as.owin(DATA$Window))
cg_ppp_10 <-ppp(x = c_g_data_10k$x, y = c_g_data_10k$y, window=as.owin(DATA$Window))

rho_e_5 <- rhohat(cg_ppp_5, DATA$Elevation) 
rho_e_8 <- rhohat(cg_ppp_8, DATA$Elevation)
rho_e_10 <- rhohat(cg_ppp_10, DATA$Elevation)

rho_f_5 <- rhohat(cg_ppp_5, DATA$Forest) 
rho_f_8 <- rhohat(cg_ppp_8, DATA$Forest)
rho_f_10 <- rhohat(cg_ppp_10, DATA$Forest)

rho_h_5 <- rhohat(cg_ppp_5, DATA$HFI) 
rho_h_8 <- rhohat(cg_ppp_8, DATA$HFI)
rho_h_10 <- rhohat(cg_ppp_10, DATA$HFI)

rho_w_5 <- rhohat(cg_ppp_5, DATA$Dist_Water) 
rho_w_8 <- rhohat(cg_ppp_8, DATA$Dist_Water)
rho_w_10 <- rhohat(cg_ppp_10, DATA$Dist_Water)

print(paste("The average rho value (elevation) of datasets with 5k, 8k and 10k rows are :",rho_e_5$ave[1]," , ",rho_e_8$ave[1]," , ",rho_e_10$ave[1]))

print(paste("The average rho value (forest) of datasets with 5k, 8k and 10k rows are :",rho_f_5$ave[1]," , ",rho_f_8$ave[1]," , ",rho_f_10$ave[1]))

print(paste("The average rho value (HFI) of datasets with 5k, 8k and 10k rows are :",rho_h_5$ave[1]," , ",rho_h_8$ave[1]," , ",rho_h_10$ave[1]))

print(paste("The average rho value (Dist_Water) of datasets with 5k, 8k and 10k rows are :",rho_w_5$ave[1]," , ",rho_w_8$ave[1]," , ",rho_w_10$ave[1]))
```
### 5. Testing if the thinned data is good representation of original data
#### 5.2 Checking if all rho value are inbetween 95% CI
```{r}
print("Elevation Rho")
quantiles_e_5 <- quantile(rho_e_5$rho, probs = c(0.975, 0.025))
quantiles_e_8 <- quantile(rho_e_8$rho, probs = c(0.975, 0.025))
quantiles_e_10 <- quantile(rho_e_10$rho, probs = c(0.975, 0.025))
if(rho_e_5$ave[1] <= quantiles_e_5[1] & rho_e_5$ave[1] >= quantiles_e_5[2]){print("5k dataset : Avaerage rho value is within 95% CI")}
if(rho_e_8$ave[1] <= quantiles_e_8[1] & rho_e_8$ave[1] >= quantiles_e_8[2]){print("8k dataset :Avaerage rho value is within 95% CI")}
if(rho_e_10$ave[1] <= quantiles_e_10[1] & rho_e_10$ave[1] >= quantiles_e_10[2]){print("10k dataset: Avaerage rho value is within 95% CI")}

print("Forest Rho")
quantiles_f_5 <- quantile(rho_f_5$rho, probs = c(0.975, 0.025))
quantiles_f_8 <- quantile(rho_f_8$rho, probs = c(0.975, 0.025))
quantiles_f_10 <- quantile(rho_f_10$rho, probs = c(0.975, 0.025))
if(rho_f_5$ave[1] <= quantiles_f_5[1] & rho_f_5$ave[1] >= quantiles_f_5[2]){print("5k dataset : Avaerage rho value is within 95% CI")}
if(rho_f_8$ave[1] <= quantiles_f_8[1] & rho_f_8$ave[1] >= quantiles_f_8[2]){print("8k dataset :Avaerage rho value is within 95% CI")}
if(rho_f_10$ave[1] <= quantiles_f_10[1] & rho_f_10$ave[1] >= quantiles_f_10[2]){print("10k dataset:Avaerage rho value is within 95% CI")}

print("HCI Rho")
quantiles_h_5 <- quantile(rho_h_5$rho, probs = c(0.975, 0.025))
quantiles_h_8 <- quantile(rho_h_8$rho, probs = c(0.975, 0.025))
quantiles_h_10 <- quantile(rho_h_10$rho, probs = c(0.975, 0.025))
if(rho_h_5$ave[1] <= quantiles_h_5[1] & rho_h_5$ave[1] >= quantiles_h_5[2]){print("5k dataset : Avaerage rho value is within 95% CI")}
if(rho_h_8$ave[1] <= quantiles_h_8[1] & rho_h_8$ave[1] >= quantiles_h_8[2]){print("8k dataset :Avaerage rho value is within 95% CI")}
if(rho_h_10$ave[1] <= quantiles_h_10[1] & rho_h_10$ave[1] >= quantiles_h_10[2]){print("10k dataset:Avaerage rho value is within 95% CI")}

print("Water Rho")
quantiles_w_5 <- quantile(rho_w_5$rho, probs = c(0.975, 0.025))
quantiles_w_8 <- quantile(rho_w_8$rho, probs = c(0.975, 0.025))
quantiles_w_10 <- quantile(rho_w_10$rho, probs = c(0.975, 0.025))
if(rho_w_5$ave[1] <= quantiles_w_5[1] & rho_w_5$ave[1] >= quantiles_w_5[2]){print("5k dataset : Avaerage rho value is within 95% CI")}
if(rho_w_8$ave[1] <= quantiles_w_8[1] & rho_w_8$ave[1] >= quantiles_w_8[2]){print("8k dataset :Avaerage rho value is within 95% CI")}
if(rho_w_10$ave[1] <= quantiles_w_10[1] & rho_w_10$ave[1] >= quantiles_w_10[2]){print("10k dataset:Avaerage rho value is within 95% CI")}
```


### 6. Conversion to ppp
```{r fig.width = 8, fig.height = 6}
cg_ppp <-ppp(x = c_g_data_10k$x, y = c_g_data_10k$y, window=as.owin(DATA$Window))
plot(cg_ppp,cols = c("maroon"))
```

### 7. Plotting Canadian Goose Locations along with elevation detail in 2D and 3D
```{r fig.width = 8, fig.height = 6}
library(viridisLite)

# Visualizing location of Canadian Goose with elevation
plot(DATA$Elevation, main = "Elevation Image")
points(cg_ppp, pch = 16, cex = 0.7, col = "white")
points(cg_ppp, pch = 16, cex = 0.3, col = "black", xpd = TRUE)

colors <- plasma(5)
# Visualizing location of Canadian Goose with elevation in 3D
fig <- persp(DATA$Elevation, 
      main = "Perspective Plot of Elevation",  
      theta = -5, phi = 45,  
      expand = 50,  
      colmap = colors, 
      border = NA,  
      apron = TRUE,
      axes = TRUE,  
      xlab = "X-axis", 
      ylab = "Y-axis", 
      zlab = "Elevation (m)",  
      ticktype = "detailed",  
      shade = 0.2,
      lwd = .5,  
      lty = "solid",  
      box = FALSE,
      visible = TRUE
)

perspPoints(cg_ppp, Z = DATA$Elevation, M = fig, pch = 16, cex = 0.4,col=rgb(1,1,1, alpha = 0.7))
```

### 8. Calculating and Visulaizing number of goose at each level of elevation
```{r}
cg_ppp_df <- as.data.frame(cg_ppp)
for (i in 1:nrow(cg_ppp_df)) {
if(length(DATA$Elevation[cg_ppp_df[i,]]) == 0){
  cg_ppp_df$value[i] <- NA
}else{
    cg_ppp_df$value[i] <- DATA$Elevation[cg_ppp_df[i,]]
  }
}
cg_ppp_df <- na.omit(cg_ppp_df)
cg_ppp_df$group <- cut(cg_ppp_df$value, breaks = 5, labels = c("Low", "Medium-Low", "Medium", "Medium-High", "High"))
elevation_vice_counts <- as.data.frame(table(cg_ppp_df$group))
ggplot(data=elevation_vice_counts, aes(x=Var1, y=Freq)) + geom_bar(stat="identity") +   labs(x = "Elevation Group", y = "Number of Geese", title = "Distribution of geese in accordance with elevation")
```

### 9. Testing homogeneity
```{r fig.width = 8, fig.height = 6}

Q <- quadratcount(cg_ppp,nx = 10, ny = 10)
plot(cg_ppp, pch = 16, cex = 0.5, cols = "darkorange",main="BC Park Location")
plot(Q, cex = 1, col = "black", add = T)
plot(intensity(Q, image = T), main = "Canadian Goose Location")
plot(cg_ppp,  pch = 16, cex = 1,cols = "grey",add = T)
quadrat.test(Q)

plot(density(cg_ppp))
```

### 10. calculating the intensity of Canadian goose locations with respect to elevation
```{r}
rho <- rhohat(cg_ppp, DATA$Elevation)
print(rho)
plot(rho,xlim = c(0, 2000), main="Relationship between intensity of Canadian Goose and elevation")
```

### 11. Calculating correlation of Canadian goose locations with respect to elevation
```{r}
lambda_park <- density(cg_ppp, bw.ppl)
Kinhom_park <- Kinhom(cg_ppp, lambda_park)
lambda_ppp_pos <- density(cg_ppp, sigma=bw.ppl, positive=TRUE)
E_park_inhom <- envelope(cg_ppp, Kinhom, simulate = expression(rpoispp(lambda_ppp_pos)),correction="border", rank = 1, nsim = 19, fix.n = TRUE)
plot(E_park_inhom, main = "K plot for inhomogenous assumption", lwd = 2)
#No correlation
```

### 12. Calculating presence of clustering
```{r}
lambda_ppp_pos <- density(cg_ppp, sigma=bw.ppl, positive=TRUE)
E_cg_inhom <- envelope(cg_ppp, pcfinhom, simulate = expression(rpoispp(lambda_ppp_pos)), rank = 1, nsim = 19)
plot(E_cg_inhom,main = "Inhomogenous Assumption envelope")
#No clustering
```
### 13. Elevation only Linear Model
```{r}
elevation_formula <- cg_ppp~Elevation
linear_elevation_fit <- ppm(elevation_formula, data=DATA)
linear_elevation_fit
plot(linear_elevation_fit, se = FALSE, superimpose = FALSE)
plot(cg_ppp, pch = 16, cex = 0.3, cols = "white", add = TRUE)
plot(cg_ppp, pch = 16, cex = 0.1, cols = "black", add = TRUE)
```
### 14. Elevation only non-linear model
```{r}
elevation_formula <- cg_ppp~Elevation+I(Elevation^2)
nonlinear_elevation_fit <- ppm(elevation_formula, data=DATA)
nonlinear_elevation_fit
plot(nonlinear_elevation_fit, se = FALSE, superimpose = FALSE)
plot(cg_ppp, pch = 16, cex = 0.3, cols = "white", add = TRUE)
plot(cg_ppp, pch = 16, cex = 0.1, cols = "black", add = TRUE)
```
###15. Elevation only Linear Model vs Elevation only non-linear model

```{r}
AIC(nonlinear_elevation_fit)
AIC(linear_elevation_fit)
anova(nonlinear_elevation_fit,linear_elevation_fit, test = "LRT")
```

###16. Checking if data deviate significantly from the Elevation only non-linear model prediction

```{r}
quadrat.test(nonlinear_elevation_fit, nx = 10, ny = 10)
#The small p value tells us that there’s a significant deviation from our model’s predictions.
```
### 17. Visualising Residual Elevation only non-linear model prediction
```{r}
library(sp)
library(sf)
res <- residuals(nonlinear_elevation_fit,type="pearson")
#res[is.infinite(res)] <- NA
#res <- na.omit(res)  # Remove missing values
#sp::plot(res,cols = "transparent")
res
```

### 18. Correlation between co-variates
```{r}
library(spatstat)
covariates_subset <- DATA[c("Elevation", "Forest", "HFI", "Dist_Water")]
cor.im(covariates_subset, use="pairwise.complete.obs")
```
### 19.checking all rho values of co-variates
```{r}
DATA$cg_ppp <-cg_ppp
elev_rho <- rhohat(cg_ppp, DATA$Elevation)
forest_rho <- rhohat(cg_ppp, DATA$Forest)
hfi_rho <- rhohat(cg_ppp, DATA$HFI)
water_rho <- rhohat(cg_ppp, DATA$Dist_Water)
par(mfrow=c(2,2)) 
plot(elev_rho,xlim = c(0, 2000), main="intensity vs elevation")
plot(forest_rho, main="intensity vs forest")
plot(hfi_rho, main="intensity vs hfi")
plot(water_rho, main="intensity vs water")
```
### 20. Modelling with all co-variates
```{r}
elevation_formula <- cg_ppp~Elevation+I(Elevation^2)+Forest+I(Forest^2)+HFI+I(HFI^2)+Dist_Water+I(Dist_Water^2)
all_covariate_fit <- ppm(elevation_formula, data=DATA)
summary(all_covariate_fit)
plot(nonlinear_elevation_fit, se = FALSE, superimpose = FALSE)
plot(cg_ppp, pch = 16, cex = 0.3, cols = "white", add = TRUE, alpha =0.3)
plot(cg_ppp, pch = 16, cex = 0.1, cols = "black", add = TRUE, alpha =0.3)
```
### 21. Comparing All Covariate with null model and Elevation only Model
```{r}
null_fit <- ppm(cg_ppp,data = DATA)

AIC(null_fit)
AIC(nonlinear_elevation_fit)
AIC(all_covariate_fit)
quadrat.test(all_covariate_fit, nx = 10, ny = 10)
anova(nonlinear_elevation_fit,null_fit, test = "LRT")
anova(all_covariate_fit,nonlinear_elevation_fit, test = "LRT")
```

## 22. Partial Residual of all co-variate model
```{r}
par_res_elev <- parres(all_covariate_fit, "Elevation")
par_res_hfi <- parres(all_covariate_fit, "HFI")
par_res_dw <- parres(all_covariate_fit, "Dist_Water")
par_res_forest <- parres(all_covariate_fit, "Forest")

#Side by side plotting
#par(mfrow = c(2,2))
plot(par_res_elev,
     legend = FALSE,
     lwd = 2,
     main = "",
     xlab = "Elevation (m)")
plot(par_res_hfi,
     legend = FALSE,
     lwd = 2,
     main = "",
     xlab = "HFI")
plot(par_res_dw,
     legend = FALSE,
     lwd = 2,
     main = "",
     xlab = "Distance from water")
plot(par_res_forest,
     legend = FALSE,
     lwd = 2,
     main = "",
     xlab = "Forest")
```

## 23. Fitting Spline Model
```{r}
library(splines)

#Fit the PPP model
fit_smooth <- ppm(cg_ppp ~ bs(Elevation, 7) + bs(Forest, 19) + bs(Dist_Water, 7) + bs(HFI, 8), data = DATA, use.gam = TRUE)

fit_smooth
```
## 24 Comparing smooth vs all- covariate model
```{r}
AIC(nonlinear_elevation_fit)
AIC(fit_smooth)
anova(nonlinear_elevation_fit,fit_smooth, test = "LRT")
```

## 25 Calculating partial residual of spline model
```{r}
glm_par_res_elev <- parres(fit_smooth, "Elevation")
glm_par_res_hfi <- parres(fit_smooth, "HFI")
glm_par_res_dw <- parres(fit_smooth, "Dist_Water")
glm_par_res_forest <- parres(fit_smooth, "Forest")

#Side by side plotting
par(mfrow = c(2,2))
plot(glm_par_res_elev,
     legend = FALSE,
     lwd = 2,
     main = "",
     xlab = "Elevation (m)")
plot(glm_par_res_hfi,
     legend = FALSE,
     lwd = 2,
     main = "",
     xlab = "HFI")
plot(glm_par_res_dw,
     legend = FALSE,
     lwd = 2,
     main = "",
     xlab = "Distance from water")
plot(glm_par_res_forest,
     legend = FALSE,
     lwd = 2,
     main = "",
     xlab = "Forest")
```

